---
title: "Monthly Sales Report"
date: today
format:
  html:
    toc: true
    embed-resources: true
  pdf:
    toc: true
---

```{r}
#| label: setup
#| include: false
orders <- read.csv("data/orders.csv", stringsAsFactors = FALSE)
orders$revenue <- orders$quantity * orders$unit_price
orders$month <- substr(orders$order_date, 1, 7)

monthly <- aggregate(
  cbind(revenue, quantity) ~ month,
  data = orders,
  FUN = sum
)
orders_count <- table(orders$month)
monthly$orders <- as.integer(orders_count[monthly$month])
monthly <- monthly[order(monthly$month), ]

best_month <- monthly$month[which.max(monthly$revenue)]
best_revenue <- max(monthly$revenue)
total_revenue <- sum(monthly$revenue)

fmt_usd <- function(x) paste0("$", formatC(x, format = "f", digits = 2, big.mark = ","))
```

## Summary

- **Total revenue:** `r fmt_usd(total_revenue)` across `r nrow(orders)` orders
- **Best month:** `r best_month` with `r fmt_usd(best_revenue)` in revenue (`r monthly$orders[monthly$month == best_month]` orders)
- **Period covered:** `r min(monthly$month)` to `r max(monthly$month)`

## Revenue by Month

```{r}
#| label: table
#| echo: false
display <- data.frame(
  Month = monthly$month,
  Orders = monthly$orders,
  Revenue = fmt_usd(monthly$revenue)
)
knitr::kable(display, align = c("l", "r", "r"))
```

Revenue climbs almost every month from **2024-07** (\$44,576.74) through the
**2024-12** peak (\$97,138.60) — more than double over six months — before
dropping sharply to \$55,698.25 in **2025-01**. From there it recovers
steadily, closing the period at \$80,435.87 in **2025-06**, just shy of the
December high. Order counts track revenue almost one-for-one (e.g. 270 orders
in July vs. 561 in December), indicating the swings are driven mainly by
*how many* orders are placed each month rather than by changes in what
customers spend per order.

## Revenue Trend

```{r}
#| label: chart
#| echo: false
#| fig-width: 8
#| fig-height: 4.5
bar_colors <- ifelse(monthly$month == best_month, "#d62728", "#1f77b4")
barplot(
  monthly$revenue,
  names.arg = monthly$month,
  las = 2,
  col = bar_colors,
  main = "Monthly Revenue",
  ylab = "Revenue (USD)",
  cex.names = 0.8
)
```

The chart makes the shape of the year obvious at a glance: a smooth ramp-up
through the second half of 2024, a single red bar marking the **2024-12**
peak (highlighted in the plot), and then a one-month cliff in January 2025
before revenue resumes climbing through the first half of 2025. No other
month comes close to matching the post-holiday drop in either direction —
the rest of the series moves in comparatively small, gradual steps.

## Discussion

```{r}
#| label: aov
#| include: false
monthly$aov <- monthly$revenue / monthly$orders
h2_2024 <- sum(monthly$revenue[monthly$month >= "2024-07" & monthly$month <= "2024-12"])
h1_2025 <- sum(monthly$revenue[monthly$month >= "2025-01" & monthly$month <= "2025-06"])
half_growth <- (h1_2025 / h2_2024 - 1) * 100
jan_drop <- (monthly$revenue[monthly$month == "2025-01"] /
             monthly$revenue[monthly$month == "2024-12"] - 1) * 100
```

**Seasonality, not pricing, drives the peak.** Average order value (AOV)
stays in a narrow \$165–\$192 band all year (`r round(min(monthly$aov),0)`
to `r round(max(monthly$aov),0)` USD), while order counts range from 270 to
561 — roughly double. December's record revenue is therefore explained
almost entirely by order *volume* (561 orders, the most of any month), not
by customers spending more per order. This is consistent with a holiday
shopping effect rather than a pricing or promotion change.

**The January drop is real but short-lived.** Revenue falls
`r round(jan_drop, 1)`% from December to January — the only double-digit
month-over-month decline in the dataset — then order volume and revenue
both recover steadily over the following five months. This pattern matches
typical post-holiday demand pullback rather than a sustained downturn: by
**2025-06**, revenue (\$80,435.87) is back above every pre-peak 2024 month.

**Underlying trend is positive across halves.** Total revenue for the second
half of 2024 (Jul–Dec) was \$`r formatC(h2_2024, format = "f", digits = 2, big.mark = ",")`,
versus \$`r formatC(h1_2025, format = "f", digits = 2, big.mark = ",")` for
the first half of 2025 — a `r round(half_growth, 1)`% increase despite H1
2025 having no equivalent holiday spike, suggesting the business is growing
its baseline (non-holiday) demand, not just riding a one-time seasonal
event.

**Limitations.** This analysis uses gross revenue (`quantity × unit_price`)
with no adjustment for returns, discounts, or refunds, and only twelve
months are available, so the "seasonal" interpretation of the December peak
cannot be confirmed without a second holiday cycle to compare against.

## Conclusion

Revenue grew steadily over the period, with a clear seasonal peak in **`r best_month`**,
driven by higher order volume (`r monthly$orders[monthly$month == best_month]` orders)
rather than higher spending per order. The sharp pullback in the following
month was temporary, and the underlying revenue trend across the full year
remained positive.
